---
title: "Towards Complete Causal Explanation with Expert Knowledge"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Towards Complete Causal Explanation with Expert Knowledge}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup}
library(expertOrientR)
```


First, we will randomly generate a MAG and the corresponding PAG.

The adjacency matrix follows the convention in `pcalg`:

* 0: No edge
* 1: Circle
* 2: Arrowhead
* 3: Tail

For example:

* `amat[a,b] = 2` and `amat[b,a] = 3` denotes the edge `a --> b`
* `amat[a,b] = 2` and `amat[b,a] = 1` denotes the edge `a o-> b`

We do not allow selection bias. Therefore, we cannot have edges `a --- b` or `a o-- b`.

```{r generate MAG and PAG}
seed <- 3

n <- 15
p <- 0.25
latent_fraction <- 0.1

mag_pag_object <- expertOrientR::generate_mag_pag(n, p, latent_fraction, seed=seed)
mag <- mag_pag_object$mag
pag <- mag_pag_object$pag
```


We will identify two circle edgemarks in the PAG and select those as expert knowledge and add them to the PAG

```{r select edges from MAG as expert knowledge}
num_expert_edges <- 2

# Find partially directed edges in the PAG
partial_edges_0 <- expertOrientR::get_partial_edges(pag)
pd_size_0 <- nrow(partial_edges_0)
num_edges <- nrow(which(pag != 0, arr.ind=T)) %/% 2

# Find all non-identified edges in the PAG
nonidentifed_edges <- which(pag == 1, arr.ind=T)
num_nonidentifed_edges <- nrow(nonidentifed_edges)

# Select background knowledge from MAG
expert_edges <- nonidentifed_edges[sample(num_nonidentifed_edges, num_expert_edges), ]

cat("Total non-identifed edges:", num_nonidentifed_edges, "\n")
cat("Number of expert knowledge edges:", num_expert_edges, "\n")
if (num_expert_edges == 0) {
    cat("\nNo non-identified edges in the PAG\n")
}
```

```{r add expert knowledge edges to PAG}
PAG_with_bg <- pag
for (i in seq_len(num_expert_edges)) {
    
    if (num_expert_edges == 1) {
        a <- expert_edges[1]
        b <- expert_edges[2]
    }
    else {
        a <- expert_edges[i, 1]
        b <- expert_edges[i, 2]
    }
    
    if (mag[a, b] == 3) {
        edge_type <- "*--"
    }
    else {
        edge_type <- "*->"
    }
    
    cat("Adding expert knowledge (", i, ") : Edge between", a, "and", b, "marked as", a, edge_type, b, "\n")
    PAG_with_bg[a, b] <- mag[a, b]
    
    # If we added a tail as background knowledge, mark the other as arrow
    if (mag[a, b] == 3) {
        PAG_with_bg[b, a] <- 2
        cat("Also marking arrowhead", row.names(PAG_with_bg)[a], "<--", row.names(PAG_with_bg)[b], "\n")
    }
}
```

Next, we complete the orientation rules.
When completeing the orientation rules, we are turning off:

* Rule 3 (see our paper for why this rule will never be invoked)
* Rules 5, 6, and 7 (we assume no selection bias and these rules involve selection bias)
* Rule 9 (see our paper for why this rule will never be invoked)
* Rule 11 (this rule is implemented as a subcase of Rule 4 for computational efficiency)

```{r complete rules}
rules_to_use <- c(TRUE, TRUE, FALSE, TRUE, FALSE,
                  FALSE, FALSE, TRUE, FALSE, TRUE,
                  FALSE, TRUE, TRUE)

result <- expertOrientR::complete_pag(PAG_with_bg, rules=rules_to_use, verbose=TRUE)
G <- result$pag
```

And finally, we verify the completeness conjecture.

```{r verify completeness}

conjecture_res <- verify_conjecture(pag, G, verbose=TRUE)

if (conjecture_res) {
    cat("Successfully verified conjecture\n")
} else {
    cat("Graph is not complete!\n")
}
```


